\mytitle

### 2016-11-01 pdocs/thesis/article-section-2.tex
acm tog styling of article continues

Using Newton's second law of motion $\vec{F}=m\vec{a}$ ,
$\vec{v}^{\Delta t}$ can be solved as
\begin{equation} \label{eq:impulseIntegraWithNewton}
\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
\end{equation}
\begin{equation} \label{eq:impulse}
m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J},
\end{equation}
where superscripts denote time, i.e. ${\vec{v}}^{\Delta t}=\vec{v}(\Delta t)$.
Next position can be found
by integrating the velocity.
Updates after each step can be summarized  for locations and  
for velocities respectively as follows:

@@

Friction in contacts and joint constraints can be handled in a unified way by refactoring
equation \ref{eq:eomV} as,  
\cite{erleben.thesis}
%\citet[p.~66-67]{erleben.thesis}
\begin{equation} \label{eq:eomV2}
\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
 J_{contact}^T \vec{\lambda}_{contact}
+ J_{joint}^T \vec{\lambda}_{joint}
+ \vec{f}_{ext}),
\end{equation}
where Jacobian terms $J_{joint}^T$ for joints are 
derived by taking time derivatives of the kinematic constraints.

Symbols used in Equation \ref{eq:eomV2} are summarized in Table
\ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
Figure \ref{fig:eom-joint} presents terms needed for joint processing.

@@

In Figure \ref{fig:eom-joint}, $\vec{r}_{anc}^{\,i}$ is used to define at which point
joint constraint is applied.

% pdf page 33, notation in typical ODEs
\begin{table}
\caption {Additional terms  for generalized equations of motion}
\label{tab:eom-g}
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}
\caption {Constraint parameters}
\label{tab:constraintParameters} 
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$\vec{v}$ & linear velocity & \\ \hline
$\vec{\omega}$ & angular velocity & \\ \hline
$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\vec{\lambda}$ & constraint force &  \\ \hline
$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin{figure}
\centering
\begin{tikzpicture}
\coordinate (Y) at (1,4);
\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
\draw[dashed](0,4) -- (Y);
\node at (-0.2,4) [align=right] {$f_y$};
\draw(0.25,1) -- (0.5,1) -- (0.5,2);
\node at (0.75,1.5) {$E$};
\node at (0.8,2.5) [anchor=west] {$\sigma = E \epsilon$ if $\sigma \le f_y$};
\end{tikzpicture}
\caption{Engineering stress-strain curve of ductile steel (not to scale).}
\label{fig:sscurve}
\end{figure}

@@

Maximum force acting in a direction of $\vec{r}_{anc}^{\,i} $
is product of area and yield stress as follows

\begin{equation} \label{eq:fN}
N_{max}= \int_A f_y
\end{equation}

Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
are a product of area and shear yield stress $\tau_y$ as follows
\begin{equation} \label{eq:fQ}
Q_{max}= \int_A \tau_y
\end{equation}

Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of the perpdendicular distance 
and yield stress $f_y$ as given for the moment around the $x$-axis 
and moment around the $z$-axis, respectively.

\begin{equation} \label{eq:Mx}
M_{max}^x= \int_A z f_y
\end{equation}

\begin{equation} \label{eq:Mz}
M_{max}^z= \int_A x f_y
\end{equation}

Maximum moment around $\vec{r}_{anc}^{\,i} $
is an integral of distance $d$ from the joint point
and shear yield stress $\tau_y$ as

\begin{equation} \label{eq:My}
M_{max}^y= \int_A d \tau_y
\end{equation}

@@

\begin {table}
\caption {Maximum forces and moments for 
rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
\label{tab:maxForces} 
\begin{center}
\begin{tabular}{| c| c|}
\hline
{\bf Direction} & {\bf Maximum value}  \\ \hline
maximum shear force & $0.5\, b\, h f_y$ \\ \hline
maximum normal force & $b\, h\, f_y$  \\ \hline
maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
\end{tabular}
\end{center}
\end {table}

For torque there is a closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where $b$ and $h$ are similar.
Better approximation for any given $b$ and $h$ can be obtained 
by integrating distance from the center of the joint over cross section and
multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
is shown in Figure \ref{fig:octave-mp}

>>>>

Using Newton's second law of motion $\vec{F}=m\vec{a}$ ,
$\vec{v}^{\Delta t}$ can be solved for the velocity as:
\begin{equation} \label{eq:impulseIntegraWithNewton}
\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
\end{equation}
\begin{equation} \label{eq:impulse}
m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J},
\end{equation}
where superscripts denote time, i.e. ${\vec{v}}^{\Delta t}=\vec{v}(\Delta t)$.
Next position can be found
by integrating the velocity.
Updates after each step can be summarized  for locations and  
for velocities respectively as follows:

@@

Friction in contacts and joint constraints can be handled in a unified way by refactoring
equation \ref{eq:eomV} as,  
\cite{erleben.thesis}
%\citet[p.~66-67]{erleben.thesis}
\begin{equation} \label{eq:eomV2}
\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
 J_{contact}^T \vec{\lambda}_{contact}
+ J_{joint}^T \vec{\lambda}_{joint}
+ \vec{f}_{ext}),
\end{equation}
where Jacobian terms $J_{joint}^T$ for joints are 
derived by taking time derivatives of the kinematic constraints.
Symbols used in Equation \ref{eq:eomV2} are summarized in Table
\ref{tab:eom-g} and Figure \ref{fig:eom-joint},
where $\vec{r}_{anc}^{\,i}$ is used to define at which point
joint constraint is applied.

@@

% pdf page 33, notation in typical ODEs
\begin{table}
\tbl{Additional terms  for generalized equations of motion}{
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}}
\label{tab:eom-g}
\end {table}

@@

\begin {table}
\tbl {Constraint parameters}{
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$\vec{v}$ & linear velocity & \\ \hline
$\vec{\omega}$ & angular velocity & \\ \hline
$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\vec{\lambda}$ & constraint force &  \\ \hline
$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}}
\label{tab:constraintParameters} 
\end {table}

@@

\begin{figure}
\centering
\begin{tikzpicture}
\coordinate (Y) at (1,4);
\draw[->] (0,0) -- (6,0) node[right] {\large{$\epsilon$}};
\draw[->] (0,0) -- (0,5) node[above] {\large{$\sigma$}};
\draw(0,0) -- (Y) -- (2,4) .. controls (5,5) .. (6,4);
\draw[dashed](0,4) -- (Y);
\node at (-0.2,4) [align=right] {$f_y$};
\draw(0.25,1) -- (0.5,1) -- (0.5,2);
\node at (0.75,1.5) {$E$};
\node at (0.8,2.5) [anchor=west] {$\sigma = E \epsilon$ if $\sigma \le f_y$};
\end{tikzpicture}
\caption{Engineering stress-strain curve of ductile steel (not to scale).}
\label{fig:sscurve}
\end{figure}

@@

Maximum force acting in a direction of $\vec{r}_{anc}^{\,i} $
is product of area and yield stress as follows:

\begin{equation} \label{eq:fN}
N_{max}= \int_A f_y.
\end{equation}

Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
are a product of area and shear yield stress $\tau_y$ as follows:
\begin{equation} \label{eq:fQ}
Q_{max}= \int_A \tau_y.
\end{equation}

Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
are integrals of the perpdendicular distance 
and yield stress $f_y$ as given for the moment around the $x$-axis 
and moment around the $z$-axis, respectively:

\begin{equation} \label{eq:Mx}
M_{max}^x= \int_A z f_y,
\end{equation}

\begin{equation} \label{eq:Mz}
M_{max}^z= \int_A x f_y.
\end{equation}

Maximum moment around $\vec{r}_{anc}^{\,i} $
is an integral of distance $d$ from the joint point
and shear yield stress $\tau_y$ as:

\begin{equation} \label{eq:My}
M_{max}^y= \int_A d \tau_y.
\end{equation}

@@

\begin {table}
\tbl {Maximum forces and moments for 
rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}{
\begin{tabular}{| c| c|}
\hline
{\bf Direction} & {\bf Maximum value}  \\ \hline
maximum shear force & $0.5\, b\, h f_y$ \\ \hline
maximum normal force & $b\, h\, f_y$  \\ \hline
maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
\end{tabular}}
\label{tab:maxForces} 
\end {table}

For torque there is a closed form solution only for
circular cross sections.
Given approximation is 
best suited for cases where $b$ and $h$ are similar.
Better approximation for any given $b$ and $h$ can be obtained 
by integrating distance from the center of the joint over cross section and
multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
is shown in Figure \ref{fig:octave-mp}.

### 2016-11-01 pdocs/thesis/article-section-3.tex
acm tog styling of article continues

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O) at (0,0,0);

@@

\begin{figure}[htb!]
\begin{description}
\item[velError] is calculated using velocities and external impulses of the connected bodies.  
 \begin{description}
\item[In constraint cases,] the main contibutor is the relative speed of the bodies at the joint point.
\item[In contact cases,] the main contributor is the relative speed of the bodies at the point of contact. 
Bodies are not allowed to penetrate each other. Restitution increases velError.
If contact is not penetrating velError is reduced by $penetration * timeStep$.
\end{description}
\item[posError] is provided by the constraint. It is a significant factor in designing stable constraints.
 \begin{description}
 \item[In fixed cases,] the value is about 12 times the actual position error. Factor 12 is based on the time step (60) 
 and the default value of error reduction parameter (erp), which has a value of 0.2 in this context.
 \item[In elastic plastic cases,]  the value is set to zero if the impulse would be larger the than maximum impulse or
the spring simulation cannot be done in a stable way.
 \item[In contact cases,] the value is zero if there is no penetration. For penetration cases it is 
$\frac{-penetration\, erp}{timeStep}$. In contact cases the default value for erp is 0.8.
 \end{description}
\item[rhs (c)] is calculated by velError jInv + posError jInv
\item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
 \begin{description}
\item[In constraint cases,] it is the mass of the body (6000).
\item[In contact cases,] it varies below the mass of the body.
\end{description}
\item[Impulse] is impulse applied to the body during a timestep.
 \begin{description}
\item[In constraint cases,] it is obtained from the btJointFeedback structure.
\item[In contact cases,] it is obtained by summing the applied impulses from the active manifolds.
\end{description}
\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. body drifting. 
Setting erp to 1 would in theory eliminate error in one step but in practice a value of 0.2 - 0.8 is used in most cases.
\end{description}
\caption{Internal variables used in \bullet\ in constraint solving.}
\label{fig:solVars}
\end{figure}

@@

\begin {table}[htb!]
\caption {Simulation values for the unconstrained case without continuous collision detection. 
Typical values are shown for internal contact values
as the number of contacts and the values at contact points differ.} 
\label{tab:freeBlockValues} 
\begin{center}
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  -0.003 & & & &  &-0.17 & 0 \\  \hline
0.533 &  -1.467 & & & & & -5.33 & 0 \\  \hline
0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
0.567 &  -1.511 & 5.67 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
0.583 &  -1.502 & 0.14 &-0.011 & 0.54& 2570  & 0.55 & 420 \\  \hline
0.600 &  -1.496 & -0.38&-0.002 & 0.1  & -1000& 0.38 & 0 \\  \hline
0.617 &  -1.492 &-0.44 & 0.004 & 0     & -1600& 0.22 & 0 \\  \hline
0.717 &  -1.497 &0.004-0.08  &-0.0003-0.001 &0-0.01 & 10-400 & -0.01 & 400 \\  \hline
0.817 &  -1.499 & & & & & -0.08 & 700 \\  \hline
0.917 &  -1.500 & & & & & -0.001 & 1000 \\  \hline
\end {tabular}
\end{center}
\end {table}

\begin{figure}[htb!]
\centering
\input{figs/uc-fig.tikz}
\caption{Graph for selected values for the unconstrained case without continuous collision detection.}
\label{fig:uc}
\end{figure}

Actual values for an unconstrained case with continuous collision detection (ccd) using 1.5 
as the radius of the ccd sphere and 0.001 as the ccd motion threshold
are shown in Table \ref{tab:freeBlockValuesWithCcd}. Collision is detected at time 0.550 s when
{\it velError} is  3.5 (5.34+0.167)-0.033/0.0167 and
{\it posError} is  0 as collision is detected before penetration. 
It should be noted that in general the ccd sphere should not extend the actual body as 
premature contacts are created if collision takes place in regions where the ccd sphere extends out of the actual body.

\begin {table}[htb!]
\caption {Simulation values for the unconstrained case with continuous collision detection. } 
\label{tab:freeBlockValuesWithCcd} 
\begin{center}
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  -0.003 & & & &  &-0.17 & 0 \\  \hline
0.550 &  -1.500 & 3.5 & 0.033 & 0 & 21000& -2 & 21000 \\  \hline
0.567 &  -1.500 & 2.17 & 0 & 0  &  8100 & 0.01 & 13000 \\  \hline
0.583 &  -1.500 & 0.15 & 0 & 0 & 600  & 0 & 937 \\  \hline
0.600 &  -1.500 & 0.17 & 0 & 0 & 600  & 0 & 1040 \\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Constraint parameter values for the fixed constraint} 
\label{tab:fixedBlockValues} 
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 & -0.0028 & & & 	 & -0.17 & 0 \\  \hline
0.033 & -0.0022 & -0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
0.050 & -0.0018 & -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
0.067 & -0.0014 &-0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
0.35... & 0 &-0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Selected constraint parameters for elastic-plastic constraint 2}
\label{tab:ep2Parameters} 
\begin{center}
\begin{tabular}{|l|p{10cm}|}
\hline
{\bf Parameter} & 
{\bf Description} 
\\ \hline
lowerLimit &  
minimum allowed translation or rotation
 \\  \hline
upperLimit &  
maximum allowed translation or rotation
 \\  \hline
springStiffness & elastic spring stiffness
 \\  \hline
enableSpring & defines if spring is active
 \\  \hline
springStiffnessLimited & should elastic behaviour be tuned to avoid instability
 \\  \hline
equilibriumPoint & should elastic behaviour be tuned to avoid instability
 \\  \hline
currentLimit & describes state of constraint \newline
 0: not limited \newline
 3: loLimit=hiLimit \newline 
 4: current value is between loLimit and HiLimit
 \\ \hline

\end {tabular}
\end{center}
\end {table}

Values for an elastic-plastic case are shown in Table \ref{tab:epBlockValues}.
The body drops freely during first simulation step and 
gains enough kinetic energy so that higher impulses are needed in the following steps.
This causes plastic strain during the next three steps.

\begin {table}[htb!]
\caption {Constraint parameter values for elastic-plastic constraint} 
\label{tab:epBlockValues} 
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\it velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017 & -0.0028 &-0.17 & 0 & -1000 & -0.17 & 0 & 0 \\  \hline
0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin {table}[htb!]
\caption {Constraint parameter values for elastic-plastic constraint 2} 
\label{tab:ep2BlockValues} 
\begin{center}
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} & 
{\bf Plastic strain} \\  \hline
0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
\end {tabular}
\end{center}
\end {table}

@@

Figure \ref{fig:charpy-series} shows three detailed pictures from the simulation.
The left picture shows early phase of simulation. 
In the middle picture the hammer has not yet collided with the specimen.
In the right picture the specimen is bending.

\begin{figure}[htb!]
\centering
\includegraphics[height=3cm]{figs/article-charpy-1}
\includegraphics[height=3cm]{figs/article-charpy-2}
\includegraphics[height=3cm]{figs/article-charpy-3}
\caption{Detailed pictures from simulation of a Charpy impact test}
\label{fig:charpy-series}
\end{figure}

@@

\begin {table}[htb!]
\caption {Energy loss for few timesteps}
\label{tab:ep2ts} 
\begin{center}
\begin{tabular}{| c| c|l|}
\hline
{\bf Timestep[ms]} & {\bf Energy loss [J]} & {\bf Notes} \\ \hline
 0.2 &  170 & specimen penetrates hammer  \\ \hline
 0.1 &  120 & \\ \hline
 0.05 &  150 & \\ \hline
\end {tabular}
\end{center}
\end {table}

@@

\begin{figure}[htb!]
\centering
\includegraphics[height=8cm]{figs/article-charpy}
\caption{Overview of simulation of Charpy impact test}
\label{fig:charpy}
\end{figure}

@@

\begin{figure}[htb!]
\centering
\includegraphics[height=5cm]{figs/demolisher-pre}
\includegraphics[height=5cm]{figs/demolisher-wip}
\includegraphics[height=5cm]{figs/demolisher-done}
\caption{Simulation of Demolisher scenario}
\label{fig:demolisher}
\end{figure}

>>>>

\begin{figure}
\centering
\begin{tikzpicture}
\coordinate (O) at (0,0,0);

@@

\begin{figure}
\begin{description}
\item[velError] is calculated using velocities and external impulses of the connected bodies.  
 \begin{description}
\item[In constraint cases,] the main contibutor is the relative speed of the bodies at the joint point.
\item[In contact cases,] the main contributor is the relative speed of the bodies at the point of contact. 
Bodies are not allowed to penetrate each other. Restitution increases velError.
If contact is not penetrating velError is reduced by $penetration * timeStep$.
\end{description}
\item[posError] is provided by the constraint. It is a significant factor in designing stable constraints.
 \begin{description}
 \item[In fixed cases,] the value is about 12 times the actual position error. Factor 12 is based on the time step (60) 
 and the default value of error reduction parameter (erp), which has a value of 0.2 in this context.
 \item[In elastic plastic cases,]  the value is set to zero if the impulse would be larger the than maximum impulse or
the spring simulation cannot be done in a stable way.
 \item[In contact cases,] the value is zero if there is no penetration. For penetration cases it is 
$\frac{-penetration\, erp}{timeStep}$. In contact cases the default value for erp is 0.8.
 \end{description}
\item[rhs (c)] is calculated by velError jInv + posError jInv
\item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
 \begin{description}
\item[In constraint cases,] it is the mass of the body (6000).
\item[In contact cases,] it varies below the mass of the body.
\end{description}
\item[Impulse] is impulse applied to the body during a timestep.
 \begin{description}
\item[In constraint cases,] it is obtained from the btJointFeedback structure.
\item[In contact cases,] it is obtained by summing the applied impulses from the active manifolds.
\end{description}
\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. body drifting. 
Setting erp to 1 would in theory eliminate error in one step but in practice a value of 0.2 - 0.8 is used in most cases.
\end{description}
\caption{Internal variables used in \bullet\ in constraint solving.}
\label{fig:solVars}
\end{figure}

@@

\begin {table*}
\tbl {Simulation values for the unconstrained case without continuous collision detection. 
Typical values are shown for internal contact values
as the number of contacts and the values at contact points differ.} {
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 &  -0.003 & & & &  &-0.17 & 0 \\  \hline
0.533 &  -1.467 & & & & & -5.33 & 0 \\  \hline
0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
0.567 &  -1.511 & 5.67 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
0.583 &  -1.502 & 0.14 &-0.011 & 0.54& 2570  & 0.55 & 420 \\  \hline
0.600 &  -1.496 & -0.38&-0.002 & 0.1  & -1000& 0.38 & 0 \\  \hline
0.617 &  -1.492 &-0.44 & 0.004 & 0     & -1600& 0.22 & 0 \\  \hline
0.717 &  -1.497 &0.004-0.08  &-0.0003-0.001 &0-0.01 & 10-400 & -0.01 & 400 \\  \hline
0.817 &  -1.499 & & & & & -0.08 & 700 \\  \hline
0.917 &  -1.500 & & & & & -0.001 & 1000 \\  \hline
\end {tabular}}
\label{tab:freeBlockValues} 
\end {table*}

\begin{figure}
\input{figs/uc-fig.tikz}
\caption{Graph for location, velocity, positional error and velocity error
for the unconstrained case without continuous collision detection.}
\label{fig:uc}
\end{figure}

Actual values for an unconstrained case with continuous collision detection (ccd) using 1.5 
as the radius of the ccd sphere and 0.001 as the ccd motion threshold
are shown in Table \ref{tab:freeBlockValuesWithCcd}. Collision is detected at time 0.550 s when
{\it velError} is  3.5 (5.34+0.167)-0.033/0.0167 and
{\it posError} is  zero all the time as collision is detected before penetration. 
It should be noted that in general the ccd sphere should not extend the actual body as 
premature contacts are created if collision takes place in regions where the ccd sphere extends out of the actual body.

\begin {table}
\tbl {Simulation values for the unconstrained case with continuous collision detection.}{ 
\begin{tabular}{|l| l|l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it pene-} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\ 
 &  & & {\it tration} &  &
 &  \\ 
 \hline
0.017 &  -0.003 & & &  &-0.17 & 0 \\  \hline
0.550 &  -1.500 & 3.5 & 0.033 & 21000& -2 & 21000 \\  \hline
0.567 &  -1.500 & 2.17 & 0 &  8100 & 0.01 & 13000 \\  \hline
0.583 &  -1.500 & 0.15 & 0 & 600  & 0 & 937 \\  \hline
0.600 &  -1.500 & 0.17 & 0 & 600  & 0 & 1040 \\  \hline
\end {tabular}}
\end {table}

@@

\begin{table}
\tbl {Constraint parameter values for the fixed constraint} {
\begin{tabular}{|l|l| l| l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & {\it posError} & {\it rhs} &
{\bf Velocity} & 
{\bf Impulse} \\  \hline
0.017 & -0.0028 & & & 	 & -0.17 & 0 \\  \hline
0.033 & -0.0022 & -0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
0.050 & -0.0018 & -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
0.067 & -0.0014 &-0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
0.35... & 0 &-0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
\end {tabular}}
\label{tab:fixedBlockValues} 
\end {table}

@@

\begin{table}
\tbl {Selected constraint parameters for elastic-plastic constraint 2}{
\begin{tabular}{|l|p{5cm}|}
\hline
{\bf Parameter} & 
{\bf Description} 
\\ \hline
lowerLimit &  
minimum allowed translation or rotation
 \\  \hline
upperLimit &  
maximum allowed translation or rotation
 \\  \hline
springStiffness & elastic spring stiffness
 \\  \hline
enableSpring & defines if spring is active
 \\  \hline
springStiffnessLimited & should elastic behaviour be tuned to avoid instability
 \\  \hline
equilibriumPoint & should elastic behaviour be tuned to avoid instability
 \\  \hline
currentLimit & describes state of constraint \newline
 0: not limited \newline
 3: loLimit=hiLimit \newline 
 4: current value is between loLimit and HiLimit
 \\ \hline
\end {tabular}}
\label{tab:ep2Parameters} 
\end {table}

Values for an elastic-plastic case are shown in Table \ref{tab:epBlockValues}.
The body drops freely during first simulation step and 
gains enough kinetic energy so that higher impulses are needed in the following steps.
This causes plastic strain during the next three steps. 
Positional error is reported as zero at start of simulation because force exceeds
maximum force.  At later phases positional error is not reported to avoid instability.

\begin{table}
\tbl{Constraint parameter values for elastic-plastic constraint}{ 
\begin{tabular}{|l|l| l| l|l|l|l|l|}
\hline
{\bf Time} & 
{\bf Location} &
{\it velError} & 
{\it rhs} &
{\it velocity} & 
{\bf Impulse} & 
{\bf Plastic } \\ 
& & & & & &{\bf strain}\\
\hline
0.017 & -0.0028 &-0.17 & -1000 & -0.17 & 0 & 0 \\  \hline
0.033 & -0.0046 &-0.33 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
0.050 & -0.0056 &-0.28 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
0.067 & -0.0056 &-0.22 & -1340 &  -0.001&  1330 & 0.004\\  \hline
0.083... & -0.0056  & -0.17 & -1000 &  0.0&  1000 & 0.004\\  \hline
\end {tabular}}
\label{tab:epBlockValues} 
\end {table}

@@

\begin{table}
\tbl {Constraint parameter values for elastic-plastic constraint 2} {
\begin{tabular}{|l| l| l|l|l|l|}
\hline
{\bf Location} &
{\it velError} &  {\it rhs} &
{\bf Impulse} & 
{\bf Plastic } \\ 
 & & & &{\bf strain}\\
\hline
 0 & -0.17  & -1000  & 1000 & 0 \\  \hline
\end {tabular}}
\label{tab:ep2BlockValues} 
\end {table}

@@

Figure \ref{fig:charpy-series} shows three detailed pictures from the simulation.
The first picture shows early phase of simulation. 
In the second picture the hammer has not yet collided with the specimen.
In the last picture the specimen is bending. 
After few more time steps specimen goes between anvils in one peace.
If maximum rotation is set lower than 1.9 radians, joint is deactivated.

\begin{figure}
\includegraphics[height=2.3cm]{figs/article-charpy-1}
\includegraphics[height=2.3cm]{figs/article-charpy-2}
\includegraphics[width=8.65cm]{figs/article-charpy-3}
\caption{Detailed pictures from simulation of a Charpy impact test}
\label{fig:charpy-series}
\end{figure}

@@

\begin {table}
\tbl {Energy loss for few timesteps}{
\begin{tabular}{| c| c|l|}
\hline
{\bf Timestep[ms]} & {\bf Energy loss [J]} & {\bf Notes} \\ \hline
 0.2 &  170 & specimen penetrates hammer  \\ \hline
 0.1 &  120 & \\ \hline
 0.05 &  150 & \\ \hline
\end {tabular}}
\label{tab:ep2ts} 
\end {table}

@@

\begin{figure}
\centering
\includegraphics[height=4.5cm]{figs/article-charpy}
\caption{Overview of simulation of Charpy impact test}
\label{fig:charpy}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=8.65cm]{figs/demolisher-pre}
\includegraphics[height=4.3cm]{figs/demolisher-wip}
\includegraphics[height=4.3cm]{figs/demolisher-done}
\caption{Simulation of Demolisher scenario}
\label{fig:demolisher}
\end{figure}
